Sparse Word Embeddings for Multi-domain Machine Translation

### 2019-05-01 2019/emnlp/sparse.tex
end alternate take for the intro

\section{Introduction \label{sec:introduction}}
% The problem
Owing to the development of flexible and powerful architectures based on neural networks \cite{bahdanau2014neural;Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date. One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpus, typically consisting of millions of parallel sentences as close as possible to the application domain. A lot of the recent research effort has focused on building MT systems in low data conditions for instance building multi-lingual \cite{} MTs or even building systems without any data \cite{}.

Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains. Domain Adaption (DA) in MT is an ancient issue \cite{Foster07amixture,Axelrod11domain}, which comes in various guise, and for which a number of solutions has been studied \cite{Chu18asurvey}. The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain. We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains.

Multi-domain adaptation is conceptually close to multi-lingual MT\cite{}, or more generally to multi-task learning \cite{}, and can be approached in a number of ways.

>>>>

\section{Introduction \label{sec:introduction}}
% The problem
Owing to the development of flexible and powerful architectures based on neural networks \cite{bahdanau2014neural;Vaswani17attention}, Machine Translation (MT) has made significant progress over the past years, and constitute the standard for most production engines to-date. One issue which repeatedly pops up with neural MT is the need to perform training with large parallel corpora, typically consisting of millions of parallel sentences deemed to be as close as possible to the application domain. A lot of the recent research effort has thus focused on developping MT systems in low data conditions, for instance building multi-lingual MTs enabling zero-shot translation\cite{} or even building systems without any data \cite{}.

Another important scenario for industrial MT systems is to adapt an neural system trained using a parallel data in one  domain to the peculiarity of other domains. Domain Adaption (DA) in MT is an ancient issue \cite{Foster07amixture,Axelrod11domain}, which comes in various guises, and for which a number of solutions have been studied \cite{Chu18asurvey}. The typical setting is \emph{supervised adaptation}, where a (small) amount of data of the target domain of interest is used to fine-tuned the parameters of a system trained on a large amount of texts in a source domain. We study here a different scenario, multi-domain adaptation \cite{Farajian17multidomain}, where we would like to use heterogenous data sources to train a unique system that would work well for all domains. This allows us to be both data-efficient (all data is used to train all domains) and compuationally efficient (we only need to train one domain). In our setting, we assume that the domain of test documents is known.

Multi-domain adaptation is conceptually close to multi-lingual MT\cite{}, or more generally to multi-task learning \cite{}, and can be approached in a number of ways. We adapt here the adaptation framework of \cite{Daume07frustratingly} to neural MT, in an attempt to confine the differences between domains at the lexical level, while the rest of the parameters is shared between domains. Our main hypothesis is that most of the differences between domains takes place at the lexical level, due to massive cross-domain polysemy. By contrast, the deeper layer, which arguably model more abstract phenomena, should be made shareable.
To this end, we make sure to design word embeddings in such that they will contain a generic as well as a domain-specific region. We perform experiments with three domains, two architectures and two language pairs and find that our technique can effectively build multi-domain NMT, outperforming several baselines, including domain-specific MT engines trained through fine-tuning. Our contribution is thus twofold:
\begin{itemize}
\item methods to adapt the idea of \cite{Daume07frustratingly} for two NMT architectures;
\item experimental results that demonstrate the effectiveness of this technique;
\end{itemize}
\fyTodo{can we train in random order ? can we get away with catastrophic forgetting ?}
\fyTodo{how to analyze the embeddings ? how can we test or claim ?}
